```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#knitr::opts_knit$set(root.dir = "/tmp") 
```

```{r libraries-required, message=FALSE, warning=FALSE}
library(tidyverse)  # data manipulation and visualization
library(modelr)     # provides easy pipeline modeling functions
library(broom)      # helps to tidy up model outputs
library(VIM)      # missing value analysis
library(mice)     # missing value analysis
library(Hmisc)    # replace missing values with mean
```


## Data loading
```{r data-load}
Kidney <- read.csv("C:/Users/Lucas/Desktop/School/Spring2021/DSC341/Final Project/Chronic_Kidney_Disease/kidney.csv")

```
## Data Cleaning

I'm going to start the cleaning process by re-coding missing values from "?" to "NA"
```{r missing-recode}
Kidney[Kidney=="?"]=NA
```


# Convert Columns to ints that need to be
```{r to-int}

Kidney$age <- as.numeric(Kidney$age)
Kidney$bp <- as.numeric(Kidney$bp)
Kidney$sg <- as.numeric(Kidney$sg)
Kidney$al <- as.numeric(Kidney$al)
Kidney$su <- as.numeric(Kidney$su)
Kidney$bgr <- as.numeric(Kidney$bgr)
Kidney$bu <- as.numeric(Kidney$bu)
Kidney$sc <- as.numeric(Kidney$sc)
Kidney$sod <- as.numeric(Kidney$sod)
Kidney$pot <- as.numeric(Kidney$pot)
Kidney$hemo <- as.numeric(Kidney$hemo)
Kidney$pcv <- as.numeric(Kidney$pcv)
Kidney$wbcc <- as.numeric(Kidney$wbcc)
Kidney$rbcc <- as.numeric(Kidney$rbcc)
```



# Missing Values Analysis

```{r missing-vals}

# Find total amount of missing cells
mean(is.na(Kidney)) * 100

# Display percentage of missing values per column (variable)
colMeans(is.na(Kidney)) * 100

# Display the number of missing values per column (variable)
colSums(is.na(Kidney))

```


# Redefine labels for categorical variables 

This adds columns to the end of the table. Each new column is named after an existing categorical variable with an "f" for "float" added to the end. (i.e. rbc -> rbcf) These new columns hold numerical representaions of the categorical variable which allows us to do things like mean. 
```{r define-labels}

Kidney$rbcf=factor(Kidney$rbc, levels=c("normal","abnormal"), labels=c(1, 2))
Kidney$pcf=factor(Kidney$pc, levels=c("normal","abnormal"), labels=c(1, 2))
Kidney$pccf=factor(Kidney$pcc, levels=c("present", "notpresent"), labels=c(1,2))
Kidney$baf=factor(Kidney$ba, levels=c("present", "notpresent"), labels=c(1,2))
Kidney$htnf=factor(Kidney$htn, levels=c("yes", "no"), labels=c(1,2))
Kidney$dmf=factor(Kidney$dm, levels=c("yes", "no"), labels=c(1,2))
Kidney$cadf=factor(Kidney$cad, levels=c("yes", "no"), labels=c(1,2))
Kidney$appetf=factor(Kidney$appet, levels=c("good", "poor"), labels=c(1,2))
Kidney$pef=factor(Kidney$pe, levels=c("yes", "no"), labels=c(1,2))
Kidney$anef=factor(Kidney$ane, levels=c("yes", "no"), labels=c(1,2))
Kidney$classf=factor(Kidney$class, levels=c("ckd", "notckd"), labels=c(1,2))

```


# Deal with missing values

The plan here is to have 2 tables total. One will not touch the data at all; leaving the missing values alone. Another will replace the missing values with the mean or median value of each variable

Kidney: has missing values, string representations for some variables (i.e. "normal"/"abnormal")
kidneyMenas: has no missing values, uses mean for missing values that are represented as ints and mean for values that are represented as strings

```{r deal-with-missing-vales}

kidneyMeans=Kidney
kidneyMeans$imputed_age <- with(kidneyMeans, impute(age, mean))
kidneyMeans$imputed_bp <- with(kidneyMeans, impute(bp, mean))
kidneyMeans$imputed_sg <- with(kidneyMeans, impute(sg, mean))
kidneyMeans$imputed_al <- with(kidneyMeans, impute(al, mean))
kidneyMeans$imputed_su <- with(kidneyMeans, impute(su, mean))
kidneyMeans$imputed_rbc <- with(kidneyMeans, impute(rbcf, median))
kidneyMeans$imputed_pc <- with(kidneyMeans, impute(pcf, mean))
kidneyMeans$imputed_pcc <- with(kidneyMeans, impute(pccf, median))
kidneyMeans$imputed_ba <- with(kidneyMeans, impute(baf, median))
kidneyMeans$imputed_bgr <- with(kidneyMeans, impute(bgr, mean))
kidneyMeans$imputed_bu <- with(kidneyMeans, impute(bu, mean))
kidneyMeans$imputed_sc <- with(kidneyMeans, impute(sc, mean))
kidneyMeans$imputed_sod <- with(kidneyMeans, impute(sod, mean))
kidneyMeans$imputed_pot <- with(kidneyMeans, impute(pot, mean))
kidneyMeans$imputed_hemo <- with(kidneyMeans, impute(hemo, mean))
kidneyMeans$imputed_pcv <- with(kidneyMeans, impute(pcv, mean))
kidneyMeans$imputed_wbcc <- with(kidneyMeans, impute(wbcc, mean))
kidneyMeans$imputed_rbcc <- with(kidneyMeans, impute(rbcc, mean))
kidneyMeans$imputed_htn <- with(kidneyMeans, impute(htnf, median))
kidneyMeans$imputed_dm <- with(kidneyMeans, impute(dmf, median))
kidneyMeans$imputed_cad <- with(kidneyMeans, impute(cadf, median))
kidneyMeans$imputed_appet <- with(kidneyMeans, impute(appetf, median))
kidneyMeans$imputed_pe <- with(kidneyMeans, impute(pef, median))
kidneyMeans$imputed_ane <- with(kidneyMeans, impute(anef, median))


kidneyMeans = kidneyMeans[c(1, 38:61, 26)]

```

# Anaylisis of outylier and unusal values

```{r outlier-analysis}

# Shows stats for outliers

#$stats = statistic used to build boxplot {min*, q1, median
  #q3, max*} (min* and max* excluding extreme values)

#$n = sample size

#$out = the values of any data points which lie beyond the
# extremes of the whiskers

# The which(is.na ...) statement will show the ID number for the corresponding value

b=boxplot.stats(Kidney$su)
b

which(is.na(match(Kidney$su, b$out))==FALSE)


par(mfrow=c(3,4))
boxplot(Kidney$age, xlab = "Age", horizontal = TRUE)
boxplot(Kidney$bp, xlab = "Blood Pressure", horizontal = TRUE)
boxplot(Kidney$su, xlab = "Sugar", horizontal = TRUE)
boxplot(Kidney$bu, xlab = "Blood Urea", horizontal = TRUE)
boxplot(Kidney$sc, xlab = "Serum Creatinine", horizontal = TRUE)
boxplot(Kidney$sod, xlab = "Sodium", horizontal = TRUE)
boxplot(Kidney$pot, xlab = "Potassium", horizontal = TRUE)
boxplot(Kidney$hemo, xlab = "Hemoglobin", horizontal = TRUE)
boxplot(Kidney$pcv, xlab = "Packed Cell Volume", horizontal = TRUE)
boxplot(Kidney$wbcc, xlab = "White Blood Cell Count", horizontal = TRUE)
boxplot(Kidney$rbcc, xlab = "Red Blood Cell Count", horizontal = TRUE)

```

# Associations Between Variables

```{r associtaions}

pairs(~age+bp+sg+su+bgr+bu+sc+sod+pot+hemo+pcv+wbcc+rbcc, data=Kidney)

```



